Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-16 IEEE_ConferenceOnGames/new_article.tex
Intro

    \section{Introduction}
    Procedural content generation is an interesting subject on the entertainment
    industry, primarily in video game development because of the savings in time
    and human resources when generating content for different video game
    projects \cite{Yannakakis2017,YannakakisContentGeneration}. In this paper an
    evolutionary content generation system is proposed to generate levels for
    the popular Angry Birds video game, taking inspiration in concepts of
    open-ended evolution. Angry birds is a game developed by Rovio Entertainment
    in which the objective is to destroy structures placed on the map as well as
    eliminating green pig-like entities in order to obtain the highest score
    possible, the game uses a gravity and collision system that allows the
    structures to fall and be destroyed by means of damage as well as the use of
    three different construction materials \textit{ice}, \textit{wood} and
    \textit{stone} each one with a different resistance, ranging from fragile,
    normal and sturdy in that order. The game mechanic is to use a slingshot to
    shoot a certain number of birds with different abilities, to destroy both
    the structures and the objectives. A high score is achieved by doing the
    most destruction with the least birds possible 
    \cite{RovioEntertainmentCorporation2009}.
    The main objective of this project is to generate levels that can be played
    by naïve or intelligent agents as well as physical players. The levels
    themselves need to be complex in the sense that interesting structures can be
    created, but at the same time, it is possible to complete in some way
    \cite{Stephenson,Stephenson2018}.
    % What is the problem?
    Procedural content generation provides a peculiar case of study because of the
    way % Is there a reference? - MG
    content has to be made for a each particular type of game. In this use case we
    have those restrictions dictated by the game, and also those imposed by a
    competition track \cite{Renz}.
    provide an interesting The challenge for the participating algorithms is to be
    flexible on the pieces that can be used and the running time of the algorithm, as
    they are determined at runtime.   
    In order to tackle this case study, we proposed the use a genetic algorithm capable 
    of creating game levels having the objective of being both \textit{interesting} and 
    \textit{playable}, the way to test our candidate solutions will be by using a open-source
    software called Science Birds \cite{sciencebirds} which uses the same resources used on the
    original angry birds game.
    % What is learned?
    In this work we found that an evolutionary algorithm can be used to create 
    this type of content not only because it is able to create levels that are interesting 
    and diverse but because an evolutionary algorithm is capable of creating things a 
    normal human could have trouble imagining of.
    % What is learned?
    % En este trabajo se generan estructuras con mayor diversidad y por lo tanto 
    % resultan ser interesantes.
    We propose a basic Open-Ended Evolution (OEE) algorithm, a variant of genetic
    evolution algorithm where the objective differs from that of a regular genetic
    algorithm,  in this case the final objective is not to reach a final state of
    the evolution (an optimal solution) but to continue evolving and creating new
    types of entities, perhaps more complex on each new generation. Solution
    entities have the possibility to evolve not only inside the frame of their type,
    but  they can create entirely new groups \cite{Standish2003}. Taylor et al.
    \cite{Taylor2016,Taylor} defines OEE as a system capable of creating novelty
    beyond a point where nothing changes on the current group of interest rather
    than just converging to a stable or quasi-stable state.For this project, OEE is
    defined simply as a genetic evolution process with different classes of objects,
    with compositional and taxonomic relationships between them. Objects of each
    class can evolve in parallel or as in this case, independently of each other. The
    idea is based on the object oriented model, in which complex behavior can be
    model using these principles. The rest of the paper is as follows.

    %% TO DO: This needs to be adjusted to the final version. We need to use
    % \ref for the numbers - MG

    Section \ref{SoA}, contains a quick overview of some aspects used by different authors
    in order to tackle this problem, their design and basic functionality, in
    Section \ref{prop_method} the description of the current state of the project is explained, how
    %% The following sentence is not clear.  % Maybe this modofocation is more understandable ? JSH
    the candidate solutions are represented and the way we defined our fitness 
    function to determine the playability and complexity of the generated levels, 
    finally in Section \ref{conclusion} we provide our conclusions about the development of this
    paper as well as improvements that can help to refine the proposed work.

>>>>

\section{Introduction}
The video game industry is always looking for content creators to satisfy
the demand for high-quality content. Current development teams include game
designers,  artists, animators,  writers, level designers, among others.
Procedural Content Generation (PCG) is often applied to generate many
aspects of the game, from maps, levels or characters, to narrative and game
design. PCG techniques can lower the cost in both time and human resources
when generating content for different video game projects, without
compromising qualitiy \cite{Yannakakis2017,YannakakisContentGeneration}. In
this paper, we introduce an evolutionary content generation system, to
generate fun and exciting levels for the popular Angry Birds video game. We
follow an Open-Ended Evolution (OEE) approach.  OOE is a type of genetic
evolution algorithm with a different objective. In this case, the goal is
not to reach a final state of in the evolution, for instance, to reach an
optimal solution in a search space. An OEE aims to continue evolving and
creating new types of entities, perhaps more complex or novel on each new
generation. Entities evolve not only inside the frame of their species, but
they can also create entirely new species \cite{Standish2003}. Taylor et al.
\cite{Taylor2016, Taylor} define OEE as a system capable of creating novelty
beyond a point where nothing changes on the current group of interest rather
than just converging to a stable or quasi-stable state. For this project,
OEE is defined simply as a genetic algorithm with different classes of
objects, with compositional and taxonomic relationships between them.
Objects in a class can evolve in parallel or as in this case, independently
of each other. The object-oriented model follows this principles to model
complex behavior and structure \cite{levesque1986knowledge}.

The case study for this paper is to follow the rules of a level generation
competition for the Angry Birds video game developed by Rovio Entertainment.
We briefly explain the game.  The objective is to destroy structures placed
on a map as well as eliminating green pig-like characters. In order to
obtain the highest score possible, the game uses a gravity and collision
system that allows the structures to fall and be destroyed by means of
damage. The structures have  three different construction materials
\textit{ice}, \textit{wood} and \textit{stone}, each one with a different
resistance, ranging from fragile, normal and sturdy in that order. The game
mechanic is to use a slingshot to shoot a certain number of birds with
different abilities, to destroy both the structures and the objectives. A
high score is achieved by causing the most destruction throwing the least
birds possible \cite{RovioEntertainmentCorporation2009}. The main objective
of this work is to generate levels that can be played by naïve or
intelligent agents as well as human players. The levels themselves need to
be complex in the sense that interesting structures can be created, but at
the same time, remaining possible to complete in some way \cite{Stephenson,
Stephenson2018}.
% What is the problem?

One peculiarity of PCG is the richness of problems and domains in which it can
be applied.  This is because each type of game dictates different restrictions
and mechanics. In this case, we have additional restrictions, dictated by the
competition track \cite{Renz}. This case study provides an interesting challenge
for the participating algorithms. The proposed PCG algorithm has to be flexible
on the pieces it can use, and also consider the running time, as both are
determined at runtime. The aim is to develop a software capable of creating game
levels having the objective of being both \textit{interesting} and
\textit{playable}. Candidate solutions, in this case, Levels of the game, will
be evaluated by using the open-source software Science Birds \cite{sciencebirds}
which uses the same resources used on the original angry birds game. Using this
software for the evaluation is an additional restriction because we are limited
to the data collected during the simulation.  
% What is learned?
In this work, we found that an evolutionary algorithm can be used to create this
type of content, because it is not only able to create levels that are
interesting and diverse but also capable of producing levels an average human
could have trouble imagining. The rest of the paper is as follows. Section
\ref{SoA}, contains a quick overview of some aspects used by different authors
to tackle this problem, their design, and basic functionality. In Section
\ref{prop_method} we describe our current proposal.  How we modeled the
candidate solutions and the way we defined our fitness function to determine the
playability and complexity of the generated levels. Finally, in Section
\ref{conclusion}, we provide our conclusions as well as improvements that can
help to refine the proposed work.
